---
title: "Model Fit Troubleshooting"
author: "Simon Hailstone"
date: "10 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=FALSE, message=FALSE}
library("knitr")
library("ggplot2"); theme_set(theme_bw())
library("dplyr")
library("tidyr")
library("reshape2")
library("fingertipsR")
library("RSQLite")
library("lme4")
library("lmerTest")
library("bbmle")
library("forcats")
library("rgdal")
library("rgeos")
library("maptools")
library("broom")
library("PerformanceAnalytics") # for scatterplot matrix
library("spdep") # for poly2nb
library("ape") # for moran's I
library("arm") # to get se.ranef function
library("kableExtra") # to make tables nicer
```

# Modelling prep

## Take a copy of dataset

```{r}
load(".RDATA")

# take a copy of the main dataset to use for modelling (some variables will be altered)
coverage_data_supplemented2 <- coverage_data_supplemented

```


## Summary stats table

```{r}
ccg_summary_stats <- coverage_data_supplemented2 %>% 
  dplyr::select(OrganisationCode,
                Eligible,
                Screened,
                Coverage,
                starts_with("Ethn_"),
                starts_with("Age_"),
                IMD_2015,
                IMD_2015_Rank,
                Satisfied_Opening_Hours,
                osnrth100km,
                gp_per_1k_eligible_women,
                nurses_per_1k_eligible_women,
                FEMALE_GP_PROPORTION,
                NON_UKQ_GP_PROPORTION,
                TOTAL_PATIENTS,
                TOTAL_NURSES_FTE) %>%
  gather(key="Variable", value="Value",-OrganisationCode) %>% 
  group_by(Variable) %>% 
  summarise(
    "n"=n(),
    "mean"=mean(Value),
    "sd"=sd(Value),
    "min"=min(Value),
    "percentile_25"=quantile(Value, probs=0.25),
    "median"=median(Value),
    "percentile_75"=quantile(Value, probs=0.75),
    "max"=max(Value)
  )


summary_stats %>%
      kable(digits=2) %>%
      kable_styling()


```

## Prep Table for model building

```{r}
# convert percentage variables by dividing by 100
coverage_data_supplemented2[,9:18] <- coverage_data_supplemented2[,9:18]/100
coverage_data_supplemented2[,10:18] <- as.vector(coverage_data_supplemented2[,10:18])
coverage_data_supplemented2$Satisfied_Opening_Hours <- as.vector(coverage_data_supplemented2$Satisfied_Opening_Hours/100)

coverage_data_supplemented2$CCG16CD <- factor(coverage_data_supplemented2$CCG16CD)
```


# Eligible Population

## Simple model with just Eligible population and no multilevel structure

Some plots first

```{r}
hist(coverage_data_supplemented2$Screened)
hist(coverage_data_supplemented2$Eligible)
hist(coverage_data_supplemented2$Coverage)


plot(coverage_data_supplemented2$Eligible,
     coverage_data_supplemented2$Screened)


plot(coverage_data_supplemented2$Eligible,
     coverage_data_supplemented2$Coverage)

```



```{r}
# no issues with normal glm
summary(glm(Coverage ~ Eligible,
               family=binomial, weights=Eligible,
               data=coverage_data_supplemented2))

```

## GLMM model with just Eligible popuation

```{r}

# Model fails to converge, also flags that model is nearly unidentifiable
glmer(Coverage ~ Eligible +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2)

```


```{r}
# Try fitting a log transformed version of Eligible populaiton

# Log transform eligible
coverage_data_supplemented2$Eligible_log <- log(coverage_data_supplemented2$Eligible)


# Model converges but warning about model being nearly unidentifiable: ery large eigenvalue
glmer(Coverage ~ Eligible_log +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2)

```


```{r}
# Try fitting a scaled version of Eligible population

# Scale eligible
coverage_data_supplemented2$Eligible_scaled <- as.vector(scale(coverage_data_supplemented2$Eligible, center=T, scale=T))

# Model converges but warning about model being nearly unidentifiable: ery large eigenvalue
glmer(Coverage ~ Eligible_scaled +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2)

```

```{r}
# Try fitting a scaled log version of Eligible population

# Scale and log eligible
coverage_data_supplemented2$Eligible_log_scaled <- as.vector(scale(log(coverage_data_supplemented2$Eligible), center=T, scale=T))

# Model converges but warning about model being nearly unidentifiable: ery large eigenvalue
glmer(Coverage ~ Eligible_log_scaled +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2)

```

## Is the answer simply that there is no relationship??

```{r}
plot(coverage_data_supplemented2$Eligible_log_scaled,
     coverage_data_supplemented2$Coverage)

```

# Nurses per 1k eligible pop

## Simple model with just Nurses per 1k eligible population and no multilevel structure

```{r}
# Exclude 291 practices with zero nurses
coverage_data_supplemented2 %>%
  filter(nurses_per_1k_eligible_women==0) %>% 
  summarise(n())


coverage_data_supplemented2 <- coverage_data_supplemented2 %>%
  filter(nurses_per_1k_eligible_women!=0)

```

Some plots first

```{r}
hist(coverage_data_supplemented2$nurses_per_1k_eligible_women)

plot(coverage_data_supplemented2$nurses_per_1k_eligible_women,
     coverage_data_supplemented2$Screened)


plot(coverage_data_supplemented2$nurses_per_1k_eligible_women,
     coverage_data_supplemented2$Coverage)

```



```{r}
# no issues with normal glm
summary(glm(Coverage ~ nurses_per_1k_eligible_women,
               family=binomial, weights=Eligible,
               data=coverage_data_supplemented2))

```


## GLMM model with just nurses_per_1k_eligible_women

```{r}

# Model converges but flags that model is nearly unidentifiable
glmer(Coverage ~ nurses_per_1k_eligible_women +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2)
```

## Scaled

```{r}
coverage_data_supplemented2$nurses_per_1k_eligible_women_scaled <- as.vector(scale(coverage_data_supplemented2$nurses_per_1k_eligible_women,center=T, scale=T))

# Model converges but flags that model is nearly unidentifiable
glmer(Coverage ~ nurses_per_1k_eligible_women_scaled +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2)

```


## Log transformed

```{r}
# Try fitting a log transformed version of nurses_per_1k_eligible_women

# Log transform
coverage_data_supplemented2$nurses_per_1k_eligible_women_log <- log(coverage_data_supplemented2$nurses_per_1k_eligible_women)


# However this does not look like a strong relationship!!
plot(coverage_data_supplemented2$nurses_per_1k_eligible_women_log,
     coverage_data_supplemented2$Coverage)



# Model converges, no warnings
glmer(Coverage ~ nurses_per_1k_eligible_women_log +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2)

summary(glmer(Coverage ~ nurses_per_1k_eligible_women_log +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2))



```

## Will the model converge if we exclude outliers from the upper end of the distribution?

```{r}

# quite a heavy handed cut
coverage_data_supplemented2 <- coverage_data_supplemented2 %>% 
  filter(nurses_per_1k_eligible_women < 5)

# Model converges but flags that model is nearly unidentifiable
glmer(Coverage ~ nurses_per_1k_eligible_women +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2)

```


## What about scaled after the outliers are excluded?

## Scaled

```{r}
coverage_data_supplemented2$nurses_per_1k_eligible_women_scaled <- as.vector(scale(coverage_data_supplemented2$nurses_per_1k_eligible_women,center=T, scale=T))


plot(coverage_data_supplemented2$nurses_per_1k_eligible_women_scaled,
     coverage_data_supplemented2$Coverage)


plot(coverage_data_supplemented2$nurses_per_1k_eligible_women_scaled,
     coverage_data_supplemented2$Eligible)

# Model converges but flags that model is nearly unidentifiable
glmer(Coverage ~ nurses_per_1k_eligible_women_scaled +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2)

```


### But this works - adjusting scaled values to be similar to dependant var
```{r}

coverage_data_supplemented2$nurses_per_1k_eligible_women_scaled_2 <- (coverage_data_supplemented2$nurses_per_1k_eligible_women_scaled + 3) / 10

plot(coverage_data_supplemented2$nurses_per_1k_eligible_women_scaled_2,
     coverage_data_supplemented2$Coverage)


# Model converges but flags that model is nearly unidentifiable
glmer(Coverage ~ nurses_per_1k_eligible_women_scaled_2 +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2)

```


## Trying some of the tips from a cross-validated forum post on this issue:
https://stats.stackexchange.com/questions/164457/r-glmer-warnings-model-fails-to-converge-model-is-nearly-unidentifiable

# First try changing the number of iterations of the optimiser

```{r}
# Model converges but flags that model is nearly unidentifiable
glmer(Coverage ~ nurses_per_1k_eligible_women +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2,
      
      glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1000000)))

```


# IMD 2015

## Simple model with just IMD and no multilevel structure

Some plots first

```{r}
hist(coverage_data_supplemented2$IMD_2015)

plot(coverage_data_supplemented2$IMD_2015,
     coverage_data_supplemented2$Screened)


plot(coverage_data_supplemented2$IMD_2015,
     coverage_data_supplemented2$Coverage)

```



```{r}
# no issues with normal glm
summary(glm(Coverage ~ IMD_2015,
               family=binomial, weights=Eligible,
               data=coverage_data_supplemented2))

```


## GLMM model with just IMD_2015

```{r}

# Model converges but flags that model is nearly unidentifiable
glmer(Coverage ~ IMD_2015 +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2)
```


```{r}
# Try fitting a log transformed version of IMD_2015

# Log transform
coverage_data_supplemented2$IMD_2015_log <- log(coverage_data_supplemented2$IMD_2015)


hist(coverage_data_supplemented2$IMD_2015_log)

plot(coverage_data_supplemented2$IMD_2015_log,
     coverage_data_supplemented2$Coverage)



# Model converges, no warnings
glmer(Coverage ~ IMD_2015_log +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2)

summary(glmer(Coverage ~ IMD_2015_log +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2))



```



```{r}
# Try fitting a scaled version of IMD_2015

# Scale
coverage_data_supplemented2$IMD_2015_scaled <- as.vector(scale(coverage_data_supplemented2$IMD_2015, scale=T, center=T))


hist(coverage_data_supplemented2$IMD_2015_scaled)

plot(coverage_data_supplemented2$IMD_2015_scaled,
     coverage_data_supplemented2$Coverage)



# Model is nearly unidentifiable 
glmer(Coverage ~ IMD_2015_scaled +
                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2)


```


# How does this affect fitting the full model?

```{r}

# Full model converges bu only if:
# IMD_2015 is log transformed
# nurses_per_1k_eligible_women is log transformed
# Eligible population is removed (which I think is justified as no relationship is apparent)

fit <- glmer(Coverage ~
                      Ethn_Asian +
                      Ethn_Black +
                      Ethn_Mixed +
                      Ethn_Other +
                      Age_25_to_49 +
                      Age_65_and_over +
                      IMD_2015_log +
                      Satisfied_Opening_Hours +
                      osnrth100km +
                      Urban_Rural +
                      FEMALE_GP_PROPORTION +
                      NON_UKQ_GP_PROPORTION +
                      nurses_per_1k_eligible_women_scaled_2 + ## Interesting

                      (1 | CCG16CD),

               family=binomial(link=logit),
               weights=Eligible,
               data=coverage_data_supplemented2)

summary(fit)



```
